Does It Scale? Symbolic Execution of PHP Web Applications

### 2017-01-22 paper.tex
abstract

\definecolor{lightgrey}{HTML}{FCFCFF}

@@

\begin{abstract}
\textcolor{gray}{
Dynamic web applications have become widely popular and are to
a large proportion based on the scripting language PHP. Output approximation of web
applications enables a range of additional tool support as well as possibilities
for vulnerability detection. Unfortunately, recent approximation approaches have
only been evaluated for smaller systems.}
\textcolor{gray}{This paper presents an experience report about the scalability of output
approximation using symbolic execution of state-of-the-art PHP web
applications. For a symbolic execution engine extended with support for
object-oriented programming and arrays, we identified language features and
corresponding programming patterns that impede symbolic execution and limit the
scalability of this approach. Our findings include: (1) Dynamic features such as
functions and includes are prone to fail for certain programming patterns. (2)
Expressions containing elements from I/O, databases or files can heavily impede
symbolic execution. Our findings provide useful guidelines to design new
tools and also to improve the development process of statically analyzable web
applications.}
\end{abstract}

@@

For our implementation of a symbolic execution engine for PHP, Oak, we have
chosen a test-driven approach with continuous integration using regression
tests based on \sf{WordPress} and \sf{SchoolMate} (see Section
\ref{sec:experiment_setup}). For \sf{SchoolMate}, a rather straightforward and
simple system, which we chose to compare our approximation results to those of
\sf{PhpSync}, we could achieve similar output approximation results. For
\sf{WordPress} though, we did not achieve high code coverage or an accurate
approximation for \sf{WordPress} as we faced a number of obstacles hindering
symbolic execution in different ways: The following Subsection
\ref{sec:experience_report} documents our design choices we made, apparent trade-offs we recognized, and limitations
we faced on our way towards a scalable symbolic execution engine; and what we
have learned from bypassing limitations faced in spite of manual effort for
future work. Subsection \ref{sec:experience_results} compares our observations
and results to \sf{PhpSync} and presents the motivation for our empirical
evaluation of our tool symbolic execution engine in Section \ref{sec:evaluation}.

\subsection{Experience Report}\label{sec:experience_report}

@@

\subsection{Experience Results} \label{sec:experience_results}
The symbolic execution engine \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928} so far does not face the
problems we documented in Section \ref{sec:experience_report}. The evaluation of
PhpSync, i.e., the PHP systems that the tool was applied to, did neither contain object-oriented
features, nor dynamic features like dynamic includes or dynamic function
invocation with symbolic values.

@@

\begin{table*}[t]
\centering 
	\begin{tabular}{lrp{5cm}rrrr}
	\toprule
	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} &
	\textbf{\#files} & \textbf{\#Output candidates} & \textbf{\#Include
	Expressions}\\
	\midrule
	AddressBook & 8.2.5.2 & Web-based address and phone book & 51,907 & 239  & 1009
	& 186\\
	SchoolMate & 1.5.4 & School administration & 8,118 & 65  & 853 & 88\\
	TimeClock & 1.0.4 & Web-based time tracking & 20,800 & 63  & 7920 & 306\\
	WebChess & 1.0.0 & Chess server software & 5,219 & 28  & 470 & 56\\
	\midrule
	Drupal & 7.5.0 & Content Management System &  52,464 & 125  & 3569 & 749\\
	phpBB & 3.1.9 & Forum & 327,371 & 1,398  & 3606 & 206\\
	phpMyAdmin & 4.6.3 & Web-based MySQL administration & 303,582 & 871  & 7103 &
	571\\
	\midrule
	Anchor & 0.12.1 & Content Management System & 15,054 & 201 & 987 & 32\\
	Kirby & 2.4.1 & Content Management System & 51.778 & 435  & 654 & 23\\
	Automad & 0.10 & Content Management System & 13.978 & 64  & 655 & 8\\
	Monstra & 3.0.4 & Content Management System & 41.102 & 490  & 1934 & 48\\
	Nibbleblog & 4.0.5 & Content Management System & 20.376 & 328  & 1013 & 28\\
	\bottomrule
	\end{tabular}
	\caption{Corpus of twelve PHP systems. The file count includes files with a .php,
	.inc, .bit or .module extension.}
	\label{corpus}
\end{table*}

>>>>

\definecolor{lightgrey}{HTML}{FFFFFF}

@@

\begin{abstract}
Dynamic web applications have become widely popular and are to a large portion
based on the scripting language PHP. Dynamic languages like PHP challenge
static analysis, as they generate code at runtime. One approach to unfold their
staged nature is symbolic execution by approximating client page output.
Knowledge about possible output enables a range of developer tools and analyses
for vulnerability detection.

However, previous tools have only been evaluated for smaller systems and
scalability for larger systems has not been shown. We criticize the missing
evalaution of practicality of respective tools and ask whether symbolic
execution for output approximation for PHP is scalable and what limitations can
be faced. This paper presents our experiences with re-implementing and
extending a previous symbolic execution engine for PHP that we evaluate for
several large PHP systems.
We find that dynamic language features are likely to be resolved imprecisely
for symbolic values and impede symbolic execution. We explore
strategies to mitigate limitations based on previous work on dynamic symbolic
execution by manually providing concrete rather than symbolic values.
\end{abstract}

@@

For our implementation of a symbolic execution engine for PHP, Oak, we have
chosen a test-driven approach with continuous integration using regression
tests based on \sf{WordPress} and \sf{SchoolMate} (see Section
\ref{sec:experiment_setup}). For \sf{SchoolMate}, a rather straightforward and
simple system, which we chose to compare our approximation results to those of
\sf{PhpSync}, we could achieve similar output approximation results. For
\sf{WordPress} though, we did not achieve high code coverage or an accurate
approximation for \sf{WordPress} as we faced a number of obstacles hindering
symbolic execution in different ways.

\subsection{Experience Report}\label{sec:experience_report}
We document our design choices we made, apparent trade-offs we
recognized, and limitations we faced on our way towards a scalable symbolic
execution engine; and what we have learned from bypassing limitations faced in
spite of manual effort for future work.

@@

% Comparison with PhpSync
\subsection{Experience Results} \label{sec:experience_results}
The symbolic execution engine \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928} so far does not face the
problems we documented in Section \ref{sec:experience_report}. The evaluation of
PhpSync, i.e., the PHP systems that the tool was applied to, did neither contain object-oriented
features, nor dynamic features like dynamic includes or dynamic function
invocation with symbolic values.

@@

\begin{table*}[t]
\centering 
	\begin{tabular}{lrp{5cm}rrrr}
	\toprule
	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} &
	\textbf{\#files} & \textbf{\#Output candidates} & \textbf{\#Include
	Expressions}\\
	\midrule
	AddressBook & 8.2.5.2 & Web-based address and phone book & 51,907 & 239  & 1009
	& 186\\
	SchoolMate & 1.5.4 & School administration & 8,118 & 65  & 853 & 88\\
	TimeClock & 1.0.4 & Web-based time tracking & 20,800 & 63  & 7920 & 306\\
	WebChess & 1.0.0 & Chess server software & 5,219 & 28  & 470 & 56\\
	\midrule
	Drupal & 7.5.0 & Content Management System &  52,464 & 125  & 3569 & 749\\
	phpBB & 3.1.9 & Forum & 327,371 & 1,398  & 3606 & 206\\
	phpMyAdmin & 4.6.3 & Web-based MySQL administration & 303,582 & 871  & 7103 &
	571\\
	\midrule
	Anchor & 0.12.1 & Content Management System & 15,054 & 201 & 987 & 32\\
	Kirby & 2.4.1 & Content Management System & 51.778 & 435  & 654 & 23\\
	Automad & 0.10 & Content Management System & 13.978 & 64  & 655 & 8\\
	Monstra & 3.0.4 & Content Management System & 41.102 & 490  & 1934 & 48\\
	Nibbleblog & 4.0.5 & Content Management System & 20.376 & 328  & 1013 & 28\\
	\bottomrule
	\end{tabular}
	\vspace{1mm}
	\caption{Corpus of twelve PHP systems. The file count includes files with a
	\tt{.php}, \tt{.inc}, \tt{.bit} or \tt{.module} extension.}
	\label{corpus}
\end{table*}
